---
title: "Homework 2"
author: "Austin Sell"
date: "April 19, 2018"
output: pdf_document
---

```{r}
data <- read.csv("Data/congress.csv")
attach(data)
library(tseries)
library(pander)
library(forecast)
options(digits = 3)
```

```{r functions, echo=FALSE}
# ARIMA Cross-validation by rolling windows
# Adapted from Rob J Hyndman's code:
# http://robjhyndman.com/hyndsight/tscvexample/
#
# Could use further generalization, e.g. to seasonality
# Careful!  This can produce singularities using categorical covariates
arimaCV <- function(x, order, xreg, include.mean, forward=1, minper=50, seasonal=NULL) {
    require(forecast)
    if (!any(class(x)=="ts")) x <- ts(x)
    n <- length(x)
    mae <- matrix(NA, nrow=n-minper, ncol=forward)
    st <- tsp(x)[1]+(minper-2)
    for(i in 1:(n-minper)) {
        xshort <- window(x, start=st+(i-minper+1), end=st+i)
        xnext <- window(x, start=st+(i+1), end=min(n, st+(i+forward)))
        xregshort <- window(xreg, start=st+(i-minper+1), end=st+i)
        xregnext <- window(xreg, start=st+(i+1), end=min(n, st+(i+forward)))
        fit <- Arima(xshort, order=order, seasonal=seasonal, xreg=xregshort, include.mean=include.mean)
        fcast <- forecast(fit, h=length(xnext), xreg=xregnext)
        mae[i,1:length(xnext)] <- abs(fcast[['mean']]-xnext)
    }
    colMeans(mae, na.rm=TRUE)
}
```

# Question 1

## Part A

```{r}
plot(Congress, DemHouseMaj, type="l")

acf(DemHouseMaj)
pacf(DemHouseMaj)

adf.test(DemHouseMaj)
PP.test(DemHouseMaj)

pre1994 <- data[StartYear<1994,]
post1994 <- data[StartYear>1994,]
pre1994$DemHouseMaj <- pre1994$DemHouseMaj - mean(pre1994$DemHouseMaj)
post1994$DemHouseMaj <- post1994$DemHouseMaj - mean(post1994$DemHouseMaj)
new.data <- rbind(pre1994, post1994)

attach(new.data)

plot(Congress, DemHouseMaj, type="l")
acf(DemHouseMaj)
pacf(DemHouseMaj)
```

## Part B

```{r}
xcovariates <- cbind(PartisanMidterm, PartisanUnem, Coattails, Pre1994)
ar0 <- arima(DemHouseMaj, order = c(0,0,0),
             xreg = xcovariates, include.mean = TRUE)
```

Table: Model Evaluation

Model Components | AIC | $\hat{\sigma}^2$ | $N$ | $\hat{\beta}_{PM}$ | $\hat{\beta}_{PU}$ | $\hat{\beta}_{CT}$ | $\hat{\beta}_{1994}$ | 
------------------|-----|-----|---|-----|-----|-----|-----|
AR(0) | `r ar0$aic` | `r ar0$sigma2` | `r length(DemHouseMaj)` | `r ar0$coef["PartisanMidterm"]` | `r ar0$coef["PartisanUnem"]` | `r ar0$coef["Coattails"]` | `r ar0$coef["Pre1994"]` |
&nbsp; | &nbsp; | &nbsp; | &nbsp; | (`r sqrt(diag(ar0$var.coef))["PartisanMidterm"]`) | (`r sqrt(diag(ar0$var.coef))["PartisanUnem"]`) | (`r sqrt(diag(ar0$var.coef))["Coattails"]`) | (`r sqrt(diag(ar0$var.coef))["Pre1994"]`) |




## Part C

```{r}
ar1 <- arima(DemHouseMaj, order = c(1,0,0),
             xreg = xcovariates, include.mean = TRUE)

ar2 <- arima(DemHouseMaj, order = c(2,0,0),
             xreg = xcovariates, include.mean = TRUE)

ma1 <- arima(DemHouseMaj, order = c(0,0,1),
             xreg = xcovariates, include.mean = TRUE)

arma11 <- arima(DemHouseMaj, order = c(1,0,1),
             xreg = xcovariates, include.mean = TRUE)
```

Table: Model Comparison

Model Components | AIC | $\hat{\sigma}^2$ | $N$ | $\hat{\beta}_{PM}$ | $\hat{\beta}_{PU}$ | $\hat{\beta}_{CT}$ | $\hat{\beta}_{1994}$ | 
------------------|-----|-----|---|-----|-----|-----|-----|
AR(0) | `r ar0$aic` | `r ar0$sigma2`| `r length(DemHouseMaj)` | `r ar0$coef["PartisanMidterm"]` | `r ar0$coef["PartisanUnem"]` | `r ar0$coef["Coattails"]` | `r ar0$coef["Pre1994"]` |
&nbsp; | &nbsp; | &nbsp; | &nbsp; | (`r sqrt(diag(ar0$var.coef))["PartisanMidterm"]`) | (`r sqrt(diag(ar0$var.coef))["PartisanUnem"]`) | (`r sqrt(diag(ar0$var.coef))["Coattails"]`) | (`r sqrt(diag(ar0$var.coef))["Pre1994"]`) |
AR(1) | `r ar1$aic` | `r ar1$sigma2` | `r length(DemHouseMaj)` | `r ar1$coef["PartisanMidterm"]` | `r ar1$coef["PartisanUnem"]` | `r ar1$coef["Coattails"]` | `r ar1$coef["Pre1994"]` |
&nbsp; | &nbsp; | &nbsp; | &nbsp; | (`r sqrt(diag(ar1$var.coef))["PartisanMidterm"]`) | (`r sqrt(diag(ar1$var.coef))["PartisanUnem"]`) | (`r sqrt(diag(ar1$var.coef))["Coattails"]`) | (`r sqrt(diag(ar1$var.coef))["Pre1994"]`) |
AR(2) | `r ar0$aic` | `r ar2$sigma2` | `r length(DemHouseMaj)` | `r ar2$coef["PartisanMidterm"]` | `r ar2$coef["PartisanUnem"]` | `r ar2$coef["Coattails"]` | `r ar2$coef["Pre1994"]` |
&nbsp; | &nbsp; | &nbsp; | &nbsp; | (`r sqrt(diag(ar2$var.coef))["PartisanMidterm"]`) | (`r sqrt(diag(ar2$var.coef))["PartisanUnem"]`) | (`r sqrt(diag(ar2$var.coef))["Coattails"]`) | (`r sqrt(diag(ar2$var.coef))["Pre1994"]`) |
MA(1) | `r ma1$aic` | `r ma1$sigma2` | `r length(DemHouseMaj)` | `r ma1$coef["PartisanMidterm"]` | `r ma1$coef["PartisanUnem"]` | `r ma1$coef["Coattails"]` | `r ma1$coef["Pre1994"]` |
&nbsp; | &nbsp; | &nbsp; | &nbsp; | (`r sqrt(diag(ma1$var.coef))["PartisanMidterm"]`) | (`r sqrt(diag(ma1$var.coef))["PartisanUnem"]`) | (`r sqrt(diag(ma1$var.coef))["Coattails"]`) | (`r sqrt(diag(ma1$var.coef))["Pre1994"]`) |
ARMA(1,1) | `r arma11$aic` | `r arma11$sigma2` | `r length(DemHouseMaj)` | `r arma11$coef["PartisanMidterm"]` | `r arma11$coef[3]` | `r arma11$coef["Coattails"]` | `r arma11$coef["Pre1994"]` |
&nbsp; | &nbsp; | &nbsp; | &nbsp; | (`r sqrt(diag(arma11$var.coef))["PartisanMidterm"]`) | (`r sqrt(diag(arma11$var.coef))["PartisanUnem"]`) | (`r sqrt(diag(arma11$var.coef))["Coattails"]`) | (`r sqrt(diag(arma11$var.coef))["Pre1994"]`) |

## Part D

```{r}
ar0.cv <- arimaCV(DemHouseMaj, order = c(0,0,0), forward=3,
                  xreg=xcovariates, include.mean = TRUE, minper = 20)

ar1.cv <- arimaCV(DemHouseMaj, order = c(1,0,0), forward=3,
                  xreg=xcovariates, include.mean = TRUE, minper = 20)

ar2.cv <- arimaCV(DemHouseMaj, order = c(2,0,0), forward=3,
                  xreg=xcovariates, include.mean = TRUE, minper = 20)

ma1.cv <- arimaCV(DemHouseMaj, order = c(0,0,1), forward=3,
                  xreg=xcovariates, include.mean = TRUE, minper = 20)

arma11.cv <- arimaCV(DemHouseMaj, order = c(1,0,1), forward=3,
                  xreg=xcovariates, include.mean = TRUE, minper = 20)
```

Table: Model Comparison

| Model Components | AIC | RMSE | MAE$_1$ | MAE$_2$ | MAE$_3$ | Average MAE |
|------------------|-----|------|---------|---------|-------|-------------|
|AR(0) | `r ar0$aic` | `r ar0$sigma2` | `r ar0.cv[1]` | `r ar0.cv[2]` | `r ar0.cv[3]` | `r mean(ar0.cv)` |
|AR(1) | `r ar1$aic` | `r ar1$sigma2` | `r ar1.cv[1]` | `r ar1.cv[2]` | `r ar1.cv[3]` | `r mean(ar1.cv)` |
|AR(2) | `r ar2$aic` | `r ar2$sigma2` | `r ar2.cv[1]` | `r ar2.cv[2]` | `r ar2.cv[3]` | `r mean(ar2.cv)` |
|MA(1) | `r ma1$aic` | `r ma1$sigma2` | `r ma1.cv[1]` | `r ma1.cv[2]` | `r ma1.cv[3]` | `r mean(ma1.cv)` |
|ARMA(1,1) | `r arma11$aic` | `r arma11$sigma2` | `r arma11.cv[1]` | `r arma11.cv[2]` | `r arma11.cv[3]` | `r mean(arma11.cv)` |