---
title: "Homework 2"
author: "Austin Sell"
date: "May 1, 2018"
output: pdf_document
---

```{r}
data <- read.csv("Data/congress.csv")
attach(data)
library(tseries)
library(pander)
library(forecast)
library(simcf)
library(tile)
library(RColorBrewer)
options(digits = 3)
```

```{r functions, echo=FALSE}
# ARIMA Cross-validation by rolling windows
# Adapted from Rob J Hyndman's code:
# http://robjhyndman.com/hyndsight/tscvexample/
#
# Could use further generalization, e.g. to seasonality
# Careful!  This can produce singularities using categorical covariates
arimaCV <- function(x, order, xreg, include.mean, forward=1, minper=50, seasonal=NULL) {
    require(forecast)
    if (!any(class(x)=="ts")) x <- ts(x)
    n <- length(x)
    mae <- matrix(NA, nrow=n-minper, ncol=forward)
    st <- tsp(x)[1]+(minper-2)
    for(i in 1:(n-minper)) {
        xshort <- window(x, start=st+(i-minper+1), end=st+i)
        xnext <- window(x, start=st+(i+1), end=min(n, st+(i+forward)))
        xregshort <- window(xreg, start=st+(i-minper+1), end=st+i)
        xregnext <- window(xreg, start=st+(i+1), end=min(n, st+(i+forward)))
        fit <- Arima(xshort, order=order, seasonal=seasonal, xreg=xregshort, include.mean=include.mean)
        fcast <- forecast(fit, h=length(xnext), xreg=xregnext)
        mae[i,1:length(xnext)] <- abs(fcast[['mean']]-xnext)
    }
    colMeans(mae, na.rm=TRUE)
}
```

# Question 1

## Part A

```{r}
plot(Congress, DemHouseMaj, type="l")

acf(DemHouseMaj)
pacf(DemHouseMaj)

adf.test(DemHouseMaj)
PP.test(DemHouseMaj)

pre1994 <- data[StartYear<1994,]
post1994 <- data[StartYear>1994,]
pre1994$DemHouseMaj <- pre1994$DemHouseMaj - mean(pre1994$DemHouseMaj)
post1994$DemHouseMaj <- post1994$DemHouseMaj - mean(post1994$DemHouseMaj)
new.data <- rbind(pre1994, post1994)


plot(new.data$Congress, new.data$DemHouseMaj, type="l")
acf(new.data$DemHouseMaj)
pacf(new.data$DemHouseMaj)
```

## Part B

```{r}
xcovariates <- cbind(PartisanMidterm, PartisanUnem, Coattails, Pre1994)
ar0 <- arima(DemHouseMaj, order = c(0,0,0),
             xreg = xcovariates, include.mean = TRUE)
```

Table: Model Evaluation

Model Components | AIC | $\hat{\sigma}^2$ | $N$ | $\hat{\beta}_{PM}$ | $\hat{\beta}_{PU}$ | $\hat{\beta}_{CT}$ | $\hat{\beta}_{1994}$ | 
------------------|-----|-----|---|-----|-----|-----|-----|
AR(0) | `r ar0$aic` | `r ar0$sigma2` | `r length(DemHouseMaj)` | `r ar0$coef["PartisanMidterm"]` | `r ar0$coef["PartisanUnem"]` | `r ar0$coef["Coattails"]` | `r ar0$coef["Pre1994"]` |
&nbsp; | &nbsp; | &nbsp; | &nbsp; | (`r sqrt(diag(ar0$var.coef))["PartisanMidterm"]`) | (`r sqrt(diag(ar0$var.coef))["PartisanUnem"]`) | (`r sqrt(diag(ar0$var.coef))["Coattails"]`) | (`r sqrt(diag(ar0$var.coef))["Pre1994"]`) |




## Part C

```{r}
ar1 <- arima(DemHouseMaj, order = c(1,0,0),
             xreg = xcovariates, include.mean = TRUE)

ar2 <- arima(DemHouseMaj, order = c(2,0,0),
             xreg = xcovariates, include.mean = TRUE)

ma1 <- arima(DemHouseMaj, order = c(0,0,1),
             xreg = xcovariates, include.mean = TRUE)

arma11 <- arima(DemHouseMaj, order = c(1,0,1),
             xreg = xcovariates, include.mean = TRUE)
```

Table: Model Comparison

Model Components | AIC | $\hat{\sigma}^2$ | $N$ | $\hat{\beta}_{PM}$ | $\hat{\beta}_{PU}$ | $\hat{\beta}_{CT}$ | $\hat{\beta}_{1994}$ | 
------------------|-----|-----|---|-----|-----|-----|-----|
AR(0) | `r ar0$aic` | `r ar0$sigma2`| `r length(DemHouseMaj)` | `r ar0$coef["PartisanMidterm"]` | `r ar0$coef["PartisanUnem"]` | `r ar0$coef["Coattails"]` | `r ar0$coef["Pre1994"]` |
&nbsp; | &nbsp; | &nbsp; | &nbsp; | (`r sqrt(diag(ar0$var.coef))["PartisanMidterm"]`) | (`r sqrt(diag(ar0$var.coef))["PartisanUnem"]`) | (`r sqrt(diag(ar0$var.coef))["Coattails"]`) | (`r sqrt(diag(ar0$var.coef))["Pre1994"]`) |
AR(1) | `r ar1$aic` | `r ar1$sigma2` | `r length(DemHouseMaj)` | `r ar1$coef["PartisanMidterm"]` | `r ar1$coef["PartisanUnem"]` | `r ar1$coef["Coattails"]` | `r ar1$coef["Pre1994"]` |
&nbsp; | &nbsp; | &nbsp; | &nbsp; | (`r sqrt(diag(ar1$var.coef))["PartisanMidterm"]`) | (`r sqrt(diag(ar1$var.coef))["PartisanUnem"]`) | (`r sqrt(diag(ar1$var.coef))["Coattails"]`) | (`r sqrt(diag(ar1$var.coef))["Pre1994"]`) |
AR(2) | `r ar0$aic` | `r ar2$sigma2` | `r length(DemHouseMaj)` | `r ar2$coef["PartisanMidterm"]` | `r ar2$coef["PartisanUnem"]` | `r ar2$coef["Coattails"]` | `r ar2$coef["Pre1994"]` |
&nbsp; | &nbsp; | &nbsp; | &nbsp; | (`r sqrt(diag(ar2$var.coef))["PartisanMidterm"]`) | (`r sqrt(diag(ar2$var.coef))["PartisanUnem"]`) | (`r sqrt(diag(ar2$var.coef))["Coattails"]`) | (`r sqrt(diag(ar2$var.coef))["Pre1994"]`) |
MA(1) | `r ma1$aic` | `r ma1$sigma2` | `r length(DemHouseMaj)` | `r ma1$coef["PartisanMidterm"]` | `r ma1$coef["PartisanUnem"]` | `r ma1$coef["Coattails"]` | `r ma1$coef["Pre1994"]` |
&nbsp; | &nbsp; | &nbsp; | &nbsp; | (`r sqrt(diag(ma1$var.coef))["PartisanMidterm"]`) | (`r sqrt(diag(ma1$var.coef))["PartisanUnem"]`) | (`r sqrt(diag(ma1$var.coef))["Coattails"]`) | (`r sqrt(diag(ma1$var.coef))["Pre1994"]`) |
ARMA(1,1) | `r arma11$aic` | `r arma11$sigma2` | `r length(DemHouseMaj)` | `r arma11$coef["PartisanMidterm"]` | `r arma11$coef[3]` | `r arma11$coef["Coattails"]` | `r arma11$coef["Pre1994"]` |
&nbsp; | &nbsp; | &nbsp; | &nbsp; | (`r sqrt(diag(arma11$var.coef))["PartisanMidterm"]`) | (`r sqrt(diag(arma11$var.coef))["PartisanUnem"]`) | (`r sqrt(diag(arma11$var.coef))["Coattails"]`) | (`r sqrt(diag(arma11$var.coef))["Pre1994"]`) |

## Part D

```{r}
ar0.cv <- arimaCV(DemHouseMaj, order = c(0,0,0), forward=3,
                  xreg=xcovariates, include.mean = TRUE, minper = 20)

ar1.cv <- arimaCV(DemHouseMaj, order = c(1,0,0), forward=3,
                  xreg=xcovariates, include.mean = TRUE, minper = 20)

ar2.cv <- arimaCV(DemHouseMaj, order = c(2,0,0), forward=3,
                  xreg=xcovariates, include.mean = TRUE, minper = 20)

ma1.cv <- arimaCV(DemHouseMaj, order = c(0,0,1), forward=3,
                  xreg=xcovariates, include.mean = TRUE, minper = 20)

arma11.cv <- arimaCV(DemHouseMaj, order = c(1,0,1), forward=3,
                  xreg=xcovariates, include.mean = TRUE, minper = 20)

allCV <- cbind(ar0.cv, ar1.cv, ar2.cv, ma1.cv, arma11.cv)
labs <- c("AR(0)", "AR(1)", "AR(2)", "MA(1)", "ARMA(1,1)")

col <- c(brewer.pal(7, "Reds")[3:7],
         brewer.pal(8, "Blues")[3:8])

matplot(allCV, type="l", col=col, lty=1, ylab="Mean Absolute Error", xlab="Periods Forward",
        main="Cross-validation of accident deaths models", xlim=c(0.75,3.75))
text(labs, x=rep(3.5,length(labs)), y=allCV[nrow(allCV),], col=col)
```

Table: Model Comparison

| Model Components | AIC | RMSE | MAE$_1$ | MAE$_2$ | MAE$_3$ | Average MAE |
|------------------|-----|------|---------|---------|-------|-------------|
|AR(0) | `r ar0$aic` | `r ar0$sigma2` | `r ar0.cv[1]` | `r ar0.cv[2]` | `r ar0.cv[3]` | `r mean(ar0.cv)` |
|AR(1) | `r ar1$aic` | `r ar1$sigma2` | `r ar1.cv[1]` | `r ar1.cv[2]` | `r ar1.cv[3]` | `r mean(ar1.cv)` |
|AR(2) | `r ar2$aic` | `r ar2$sigma2` | `r ar2.cv[1]` | `r ar2.cv[2]` | `r ar2.cv[3]` | `r mean(ar2.cv)` |
|MA(1) | `r ma1$aic` | `r ma1$sigma2` | `r ma1.cv[1]` | `r ma1.cv[2]` | `r ma1.cv[3]` | `r mean(ma1.cv)` |
|ARMA(1,1) | `r arma11$aic` | `r arma11$sigma2` | `r arma11.cv[1]` | `r arma11.cv[2]` | `r arma11.cv[3]` | `r mean(arma11.cv)` |

AR(0) is best model

## Part E

```{r}
n.ahead <- 3
PartisanMidterm.cf <- c(-1,0,1)
Unemployment.cf1 <- rep(4.6, n.ahead)
Unemployment.cf2 <- rep(3.6, n.ahead)
Unemployment.cf3 <- rep(5.6, n.ahead)
PartisanUnem.cf1 <- c(-1,1,1)*Unemployment.cf1
PartisanUnem.cf2 <- c(-1,1,1)*Unemployment.cf2
PartisanUnem.cf3 <- c(-1,1,1)*Unemployment.cf3
Coattails.cf <- c(0,1,0)
Pre1994.cf <- rep(0, n.ahead)

cf1.cov <- as.data.frame(cbind(PartisanMidterm.cf, PartisanUnem.cf1, Coattails.cf, Pre1994.cf))
cf2.cov <- as.data.frame(cbind(PartisanMidterm.cf, PartisanUnem.cf2, Coattails.cf, Pre1994.cf))
cf3.cov <- as.data.frame(cbind(PartisanMidterm.cf, PartisanUnem.cf3, Coattails.cf, Pre1994.cf))

names(cf1.cov) <- c("PartisanMidterm", "PartisanUnem", "Coattails", "Pre1994")
names(cf2.cov) <- c("PartisanMidterm", "PartisanUnem", "Coattails", "Pre1994")
names(cf3.cov) <- c("PartisanMidterm", "PartisanUnem", "Coattails", "Pre1994")

ypred1 <- predict(ar0, n.ahead=n.ahead, newxreg = cf1.cov)
ypred2 <- predict(ar0, n.ahead=n.ahead, newxreg = cf2.cov)
ypred3 <- predict(ar0, n.ahead=n.ahead, newxreg = cf3.cov)

# Make a plot
plot.new()
par(usr = c(0, length(DemHouseMaj) + n.ahead, -50, 80) )
# make the x-axis
axis(1,
     at = seq(from = 1, to = 31, by = 1),
     labels = seq(from = 1963, to = 2023, by = 2)
)
axis(2)

title(xlab = "Starting Year of Congressional Session",
      ylab = "Size of Democratic House Majority",
      main="Predicted effect of Scenario 1")


# Polygon of predictive interval for no law (optional)
x0 <- (length(DemHouseMaj)+1):(length(DemHouseMaj) + n.ahead)
y0 <- c(ypred1$pred - 2*ypred1$se, rev(ypred1$pred + 2*ypred1$se), (ypred1$pred - 2*ypred1$se)[1] )
polygon(x = c(x0, rev(x0), x0[1]),
        y = y0,
        border=NA,
        col="#FFBFBFFF"
)

# Plot the actual data
lines(x = 1:length(DemHouseMaj),
      y = DemHouseMaj
)

# Add the predictions for no law
lines(x = length(DemHouseMaj):(length(DemHouseMaj)+n.ahead),
      y = c(DemHouseMaj[length(DemHouseMaj)],ypred1$pred),  # link up the actual data to the prediction
      col = "red"
)

# Make a plot 2
plot.new()
par(usr = c(0, length(DemHouseMaj) + n.ahead, -50, 80) )
# make the x-axis
axis(1,
     at = seq(from = 1, to = 31, by = 1),
     labels = seq(from = 1963, to = 2023, by = 2)
)
axis(2)

title(xlab = "Starting Year of Congressional Session",
      ylab = "Size of Democratic House Majority",
      main="Predicted effect of Scenario 2")


# Polygon of predictive interval for no law (optional)
x0 <- (length(DemHouseMaj)+1):(length(DemHouseMaj) + n.ahead)
y0 <- c(ypred2$pred - 2*ypred2$se, rev(ypred2$pred + 2*ypred2$se), (ypred2$pred - 2*ypred2$se)[1] )
polygon(x = c(x0, rev(x0), x0[1]),
        y = y0,
        border=NA,
        col="#FFBFBFFF"
)

# Plot the actual data
lines(x = 1:length(DemHouseMaj),
      y = DemHouseMaj
)

# Add the predictions for no law
lines(x = length(DemHouseMaj):(length(DemHouseMaj)+n.ahead),
      y = c(DemHouseMaj[length(DemHouseMaj)],ypred2$pred),  # link up the actual data to the prediction
      col = "red"
)

# Make a plot 3
plot.new()
par(usr = c(0, length(DemHouseMaj) + n.ahead, -50, 80) )
# make the x-axis
axis(1,
     at = seq(from = 1, to = 31, by = 1),
     labels = seq(from = 1963, to = 2023, by = 2)
)
axis(2)

title(xlab = "Starting Year of Congressional Session",
      ylab = "Size of Democratic House Majority",
      main="Predicted effect of Scenario 3")


# Polygon of predictive interval for no law (optional)
x0 <- (length(DemHouseMaj)+1):(length(DemHouseMaj) + n.ahead)
y0 <- c(ypred3$pred - 2*ypred3$se, rev(ypred3$pred + 2*ypred3$se), (ypred3$pred - 2*ypred3$se)[1] )
polygon(x = c(x0, rev(x0), x0[1]),
        y = y0,
        border=NA,
        col="#FFBFBFFF"
)

# Plot the actual data
lines(x = 1:length(DemHouseMaj),
      y = DemHouseMaj
)

# Add the predictions for no law
lines(x = length(DemHouseMaj):(length(DemHouseMaj)+n.ahead),
      y = c(DemHouseMaj[length(DemHouseMaj)],ypred3$pred),  # link up the actual data to the prediction
      col = "red"
)

```

Table: Counterfactual Values

| Variable | 2018 | 2020 | 2022 |
|--------------|------|------|------|
|**Scenario 1** | &nbsp; | &nbsp; | &nbsp; |
|Partisan Midterm | `r PartisanMidterm.cf[1]` | `r PartisanMidterm.cf[2]` | `r PartisanMidterm.cf[3]` |
|Partisan Unemployment | `r PartisanUnem.cf1[1]` | `r PartisanUnem.cf1[2]` | `r PartisanUnem.cf1[3]` |
|Coattails | `r Coattails.cf[1]` | `r Coattails.cf[2]` | `r Coattails.cf[3]` |
|Pre-1994 | `r Pre1994.cf[1]` | `r Pre1994.cf[2]` | `r Pre1994.cf[3]` |
|&nbsp;|&nbsp;|&nbsp;|&nbsp;|
|**Scenario 2** | &nbsp; | &nbsp; | &nbsp; |
|Partisan Midterm | `r PartisanMidterm.cf[1]` | `r PartisanMidterm.cf[2]` | `r PartisanMidterm.cf[3]` |
|Partisan Unemployment | `r PartisanUnem.cf2[1]` | `r PartisanUnem.cf2[2]` | `r PartisanUnem.cf2[3]` |
|Coattails | `r Coattails.cf[1]` | `r Coattails.cf[2]` | `r Coattails.cf[3]` |
|Pre-1994 | `r Pre1994.cf[1]` | `r Pre1994.cf[2]` | `r Pre1994.cf[3]` |
|&nbsp;|&nbsp;|&nbsp;|&nbsp;|
|**Scenario 3** | &nbsp; | &nbsp; | &nbsp; |
|Partisan Midterm | `r PartisanMidterm.cf[1]` | `r PartisanMidterm.cf[2]` | `r PartisanMidterm.cf[3]` |
|Partisan Unemployment | `r PartisanUnem.cf3[1]` | `r PartisanUnem.cf3[2]` | `r PartisanUnem.cf3[3]` |
|Coattails | `r Coattails.cf[1]` | `r Coattails.cf[2]` | `r Coattails.cf[3]` |
|Pre-1994 | `r Pre1994.cf[1]` | `r Pre1994.cf[2]` | `r Pre1994.cf[3]` |


